\chapter*{ACKs}
    % Given how the \chpater{} command is defined this line is explicitly needed
        % despite having set \pagestyle{empty} before...
        % See https://tex.stackexchange.com/questions/19738/why-doesnt-pagestyleempty-work-on-the-first-page-of-a-chapter
    \thispagestyle{empty}

    This project marks the end of a key step in my academic and personal lives. I could not have tackled such a task without the help I have been gifted by many more individuals than I can really account for. However, I will try to do my best to thank as many as I possibly can.\\

    I would like to thank my parents. Even though when asked about what I do they just say ``there are a bunch of weird letters moving on screen'' they have always been there for me along the way. Figures such as my uncle, \textit{Chema Luzón} and my grandmother, \textit{Josefina Álvarez} have also played a crucial role in my academic progression: they showed me, by example, the meaning and value of pure, raw knowledge.\\

    If there is something I know for sure is I would not have made it here if it were not for all the teachers I have had, both good and bad. The best teachers have not only transmitted the technical knowledge I have made use of time and time again, but they also taught me how to think and how to appreciate knowledge for what it is. Teachers such as \textit{Francisco José Álvarez}, \textit{Antonio García}, \textit{José Manuel Giménez}, \textit{Javier Macías}, \textit{Iván Marsá}, \textit{Isaías Martínez}, \textit{Manuel Moreno}, \textit{Rosalino Pulido} and \textit{Elisa Rojas} have given me more than they know.\\

    It is quite difficult to understate the role my friends and loved ones have played in the years leading to this moment. Classmates such as \textit{David Gaupp} and \textit{Cristian Morales}, long time friends such as \textit{Carlos Corbella}, \textit{Gonzalo Martínez} and \textit{Alberto Río} and my partner, \textit{Alicia}, have made the road a thrill.\\

    I would finally like to personally thank \textit{David Carrascal}. He has not only provided direct help on this project (including this report) and many others, but he has also been one of the people we have shared the most with. Despite meeting towards the end of the degree he has proven to be, time and time again, the most knowledgeable person I know. His drive and passion for what he does has motivated me to try and improve every single day. Nonetheless and, more importantly, he has also been a great friend who makes you feel at ease with no matter what.\\

    I am certain somebody is being left out. One of the best things of knowing nothing is that everybody can teach and show you something. Given there is not enough paper in the world to write down every single person's name I only ask that nobody feels ignored: everybody I have ever met has played a role, either big or small, in me being here. I cannot say anything but \textit{thank you}.\\
